
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load the required libraries
library(readr)

```
2. Consider the problem of understanding the risk of countries for foreign investment.
The dataset "countryrisk.csv" is attached, where for each country, four features of
corruption index, peace index, legal risk index, and real GDP growth rate are measured.

```{r load_countryriskdata}

temp_data <- read.csv("countryriskdata.csv", row.names = 1) 
countryrisk <- subset(temp_data, select = -c(Abbrev))
head(countryrisk)

```
a. Carry out K-means clustering for K = 3.
```{r 2a}
#random seed so that K-means outputs reproducible.
set.seed(8) 

#multiple run with different starting point to avoid local optimum:

km.out <- kmeans(countryrisk,centers = 3, nstart=10) #starting point = 10
km.out$tot.withinss
km.out <- kmeans(countryrisk,centers = 3, nstart=20) #starting point = 20
km.out$tot.withinss
km.out <- kmeans(countryrisk,centers = 3, nstart=50) #starting point = 50
km.out$tot.withinss

#minimum total within-cluster sum of squares when starting point = 20 or 50.
#continue with starting point = 50.
km.out

countryrisk_cluster <- cbind(countryrisk, cluster = km.out$cluster)
countryrisk_cluster[
  order(countryrisk_cluster[,"cluster"]),
  ]

```
(b) Use hierarchical clustering (with complete linkage) to determine 3 clusters.
```{r 2b}

hc.complete = hclust(dist(countryrisk), method="complete")
hc.complete

hc.clusters=cutree(hc.complete,3)
table(hc.clusters)

#visualize the cluster
plot(hc.complete, main =" Complete Linkage ", xlab="", sub ="", cex = 0.3, hang = -1)
rect.hclust(hc.complete, k = 3, border = 2:5)

```

(c-1) K-means clustering for K = 3 with scaled features.

```{r c-1}

countryrisk_scaled <- scale(countryrisk) #scale the features
head(countryrisk_scaled)

#random seed so that K-means outputs reproducible.
set.seed(8) 

#multiple run with different starting point to avoid local optimum:

km.out.scaled <- kmeans(countryrisk_scaled,centers = 3, nstart=10) #starting point = 10
km.out.scaled$tot.withinss

km.out.scaled <- kmeans(countryrisk_scaled,centers = 3, nstart=20) #starting point = 20
km.out.scaled$tot.withinss

km.out.scaled <- kmeans(countryrisk_scaled,centers = 3, nstart=50) #starting point = 50
km.out.scaled$tot.withinss

#minimum total within-cluster sum of squares when starting point = 10 or 20 or 50.
#continue with starting point = 50.
km.out.scaled

countryrisk_scaled_cluster <- cbind(countryrisk_scaled, cluster = km.out$cluster)
countryrisk_scaled_cluster[
  order(countryrisk_scaled_cluster[,"cluster"]),
  ]


```

(c-2) Use hierarchical clustering (with complete linkage) to determine 3 clusters with scaled features.
```{r c-2}

hc.complete.scaled = hclust(dist(countryrisk_scaled), method="complete")
hc.complete.scaled
hc.clusters.scaled=cutree(hc.complete.scaled,3)
table(hc.clusters.scaled)

#visualize the cluster
plot(hc.complete.scaled, main =" Complete Linkage ", xlab="", sub ="", cex = 0.3, hang = -1)
rect.hclust(hc.complete.scaled, k = 3, border = 2:5)

```

(d) For each of the clustering results in parts (a), (b), & (c), comment on the risk level (high, moderate, low) of each cluster. Compare the results (K-means clustering vs. hierarchical clustering; unscaled data vs. scaled data).

Corruption: the corruption index is on a scale from 0 (highly corrupt) to 100 (no corruption)

Peace: the peace index is on a scale from 1 (very peaceful) to 5 (not at all peaceful)

Legal: the legal risk index runs from 0 to 10 (with high values being favorable)

GDP Growth: the real GDP growth (%) is typically a positive or negative number with a magnitude less than 10. 


### K-MEANS CLUSTERING COMPARISON (SCALED VS UNSCALED)
```{r d-1}
# UNSCALED:
km.out$size
km.out$centers

# SCALED:
km.out.scaled$size
km.out.scaled$centers
```
Comment:

Using unscaled data:

- Cluster 1 (34 countries): medium risk countries - high GDP growth.

- Cluster 2 (67 countries): high risk countries (low corruption index = high corruption, high peace index = not peaceful, low legal index = not favorable) - medium GDP growth.

- Cluster 3 (21 countries): low risk countries (high corruption index = no corruption, low peace index = very peaceful, high legal index = favorable) - low GDP growth. 

Using scaled data:

- Cluster 1 (69 countries): medium risk countries - high GDP growth. 

- Cluster 2 (38 countries): low risk countries (high corruption index = no corruption, low peace index = very peaceful, high legal index = favorable) - medium GDP growth.

- Cluster 3 (15 countries): high risk countries (low corruption index = high corruption, high peace index = not peaceful, low legal index = not favorable) - low GDP growth.


### HIERARCHICAL CLUSTERING COMPARISON (SCALED VS UNSCALED)
```{r d-2}
# UNSCALED:
hc.clusters=cutree(hc.complete,3)
table(hc.clusters)
plot(hc.complete, main =" Complete Linkage ", xlab="", sub ="", cex = 0.3, hang = -1)
rect.hclust(hc.complete, k = 3, border = 2:5)

# SCALED:
hc.clusters.scaled=cutree(hc.complete.scaled,3)
table(hc.clusters.scaled)
plot(hc.complete.scaled, main =" Complete Linkage ", xlab="", sub ="", cex = 0.3, hang = -1)
rect.hclust(hc.complete.scaled, k = 3, border = 2:5)
```
Comment:

The classification is based on comparison to results from K-means clustering.
Using unscaled data:

- Cluster 1: 59 countries - high risk

- Cluster 2: 21 countries - low risk

- Cluster 3: 42 countries - medium risk 

The classification is based on comparison to results from K-means clustering.
Using scaled data:

- Cluster 1: 76 countries - medium risk

- Cluster 2: 44 countries - low risk

- Cluster 3: 2 countries - high risk


### K-MEANS VS HIERARCHICAL CLUSTERING COMPARISON
```{r d-3}
# UNSCALED
km.clusters = km.out$cluster
table(km.clusters,hc.clusters)

# SCALED
km.clusters.scaled = km.out.scaled$cluster
table(km.clusters.scaled,hc.clusters.scaled)
```
Comment:

Using unscaled data:

- All 59 countries in cluster 1 using hierarchical clustering (HC) are grouped to cluster 3 by K-means clustering (KM) - high risk countries.

- All 21 countries in cluster 2 using HC are grouped to cluster 3 using KM - low risk countries.

- Most (34 countries) in cluster 3 using HC are grouped to cluster 1 using KM, with some (8 countries) are grouped to cluster 2 using KM. 

Using scaled data:

- Most of cluster 1 (62 countries) using HC are grouped to cluster 1 using KM, but 13 countries in HC are grouped to cluster 3 in KM and 1 country in HC is grouped to clsuter 2.

- Most of cluster 2 (37 countries) using HC are grouped to cluster 2 using KM, but 7 countries in HC are grouped to cluster 1 in KM.

- All 2 countries in cluster 3 using HC are grouped to cluster 3 using KM.
