
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# load the required libraries
library(readr)

```
1. Use the attached dataset "iris.csv" to conduct the principal component analysis.

```{r load_iris}

iris <- read_csv("iris.csv")

```
(a) Delete the column containing the species type and use "newiris" for this new
dataset.

```{r 1a}

newiris <- subset(iris, select = -c(Species))
head(newiris)

```
(b) Perform a principal component analysis on the dataset newiris created above
with "scale=TRUE". Save your PCA results in "irpc". Print "irpc" to see the
output.

```{r 1b}
pca.out=prcomp(newiris,scale=TRUE) 
#prcomp() centers the variables to have mean zero by default.
#scale=TRUE scales the variables to have standard deviation one.
pca.out

irpc <- capture.output(pca.out)
irpc
```
(c) Compare output in "irpc" with "eigen(cor(newiris))". What do you notice?

```{r 1c}

eigen.out = eigen(cor(newiris))
eigen.out

#The eigenvectors produced in eigen.out results are exactly the same as PC loading vectors produced in irpc.

```
(d) Write down the first two PCs (in formula) and calculate the proportion of variance explained by PC1 and PC2, respectively according to the output in "irpc".

```{r 1d}
pca.var = pca.out$sdev^2
pca.pve = pca.var / sum (pca.var)
pca.pve

```

$$ y^{hat} _{1}   = 0.52x_{1} - 0.27x_{2} + 0.58x_{3} + 0.56x_{4} $$
PC1 = 72.96%

PC1 that accounts for 72.96% of the variance places almost equal weights on features 'Sepal_len', 'Petal_len' and 'Petal_wid'.

$$ y^{hat} _{2} = -0.38x_{1} - 0.92x_{2} - 0.02x_{3} - 0.07x_{4}$$
PC2 = 22.85%

PC2 that accounts for 22.85% of variance places most of the weight on Sepal_wid. 
It shows that feature 'Sepal_wid' provides different information than the other three features.
